Sampling requirements to detect ecosystem change

### 2020-01-15 R/manuscript_draft.Rmd
edits draft

@@

<img src="Fig_conceptual.pdf" alt="conceptual figure"  width="4200" height="6200">

@@

Data are reconstructed cladoceran assemblage from a 74-cm sediment core (VAR10-10) covering the 1816(±26)—2010 period. Half of VAR 10.10 was sliced at a 1-cm interval and was used to count cladoceran remains (Bruel et al, 2018). We defined the "true" changepoint as the one identified from the complete record, in 1983/1988. 
We used the same ordination method as in Bruel et al (2018), but note that DCA is best used to conserve multiple gradients. As a result, several changepoints were detected on the first component (in 1926/1928, 1946/1948 and 1983/1988). We conducted the analysis on the second component as well, where only one changepoint was detected (in 1944/1946).

We used the full record (74 observation) and subsampled following the three strategies described earlier (random, regular, iterative), varrying the maximum number analyzed from 5 to 15.

@@

@@

>>>>

Methods to dectect change overtime: see summary https://lindeloev.github.io/mcp/articles/packages.html

@@

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7.5}
source(paste0(getwd(),"/R/Fig_conceptual.R"))

We hypothesize  that random sampling will perform the worst at getting as close as possible to a "real" changepoint.

@@

Data are reconstructed cladoceran assemblage from a 74-cm sediment core (VAR10-10) covering the 1816(±26)—2010 period. Half of VAR 10.10 was sliced at a 1-cm interval and was used to count cladoceran remains (Bruel et al, 2018). We defined the "true" changepoint as the one identified from the complete record, in 1983/1988. 
Changepoints were detected on first component computed from the data sub-sampled. We used the same ordination method as in Bruel et al (2018), but note that DCA is best used to conserve multiple gradients. As a result, several changepoints were detected on the first component (in 1926/1928, 1946/1948 and 1983/1988). We conducted the analysis on the second component as well, where only one changepoint was detected (in 1944/1946).

We used the full record (74 observation) and subsampled following the three strategies described earlier (random, regular, iterative), varrying the maximum number analyzed from 5 to 15. Iterative sampling final number could be lower than maximum number if both samples framing the shift were analyzed.

@@

__Main point__: Changepoint was detected even though we used a transformed vector (from community to first component of DCA)

@@

Limits: Systematic sampling of a time-series is required to detect critical slowing-down or early warning of shifts (ref, maybe among others Frossard et al 2015, Doncaster et al 2016). The lower number of input data prior to ordination did not impact the finding of the main transition. However, if other metrics are of interest to the investigators, such as species richness, a more systematic or at least more extensive approach would be require.
